Hydras, Ordinals \& Co.\\ A library in Coq of entertaining formal mathematics

### 2021-10-26 doc/gaia-chapter.tex
more snippets

The module~\href{../theories/html/gaia_hydras.T1Bridge_use.html}%
{\texttt{gaia\_hydras.T1Bridge\_use}} shows an example of
application of \texttt{multA}.

@@

>>>>

The module~\href{../theories/html/gaia_hydras.T1Bridge_use.html}%
{\texttt{gaia\_hydras.T1Bridge\_use}} shows a small
example of importation of \texttt{multA} into \texttt{hydra-battles}' world.

@@

\section{Gaia's proof of well-foundedness}
The library \texttt{gaia-hydras} already contains two proofs
of well-foundedness of the strict order on ordinal terms in
Cantor normal form: a direct proof~\vref{sec:T1Wf} and a proof based on the recursive path ordering~\vref{remark:a3pat}, by
\'Evelyne Contejean. We found it interesting to comment Gaia's proof, which uses  tools already present in~\cite{CantorContrib}.

\subsection{Restricted recursion}
As remarked in~\vref{sec:T1Wf}, the order \texttt{T1lt} on \texttt{T1} is not well-founded, because of terms not in normal form.
But the \emph{restriction} of \texttt{T1lt} to terms in normal form \emph{is} well-founded. The following section introduces a  vocabulary in order to reason about statements of the form  ``the restriction of the relation $R$ to a subset $P$ of $A$ is well founded''.

\inputsnippets{nfwfgaia/restrictedRecursion}

The following induction principle is just a variant of \texttt{well\_founded\_induction\_type} for relations defined with \texttt{restrict}.

\inputsnippets{nfwfgaia/restrictedRecursiona}

With $P=\texttt{T1nf}$, this principle will allow us to write proofs by transfinite induction.

\subsection{Main proof steps}

\emph{Note: the abreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.}

\inputsnippets{nfwfgaia/nfWfProofa}

The following elimination on \texttt{a} pushes a new hypothesis \texttt{Hb} which --- in the same terms as in Section~\vref{sec:strongly-accessible} --- tells that any ordinal less than $a$ is ``strongly accessible''.

\inputsnippets{nfwfgaia/nfWfProofb, nfwfgaia/nfWfProofc}

Now, let us consider any ordinal term which is (by \texttt{LT}) less than \texttt{(cons a n b)}.

There are three cases to consider:
\begin{itemize}
\item It is of the form \texttt{(cons a' n' b')} where \texttt{a'} is strictly less than \texttt{a}. Then, it is accessible (by hypothesis \texttt{Hb}).
  \inputsnippets{nfwfgaia/nfWfProofd}

\end{itemize}

### 2021-10-26 doc/part-hydras.tex
small change in a snippet
fix typos
more snippets

\subsection{About  well-foundedness}
\label{sec:orgheadline82}
   In order to use \texttt{T1} for proving termination results,
we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.

@@

\subsubsection{Using a stronger inductive predicate.}
\label{sec:orgheadline78}
  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
through \texttt{LT}, we propose to show first that any well formed 
term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).

@@

\item Finally, $\canonseq{0}{i}= \alpha$.
\end{itemize}
\end{mathframe}

@@

\input{movies/snippets/Hprime/HprimeDef}

@@

In order to make this statement more readable, we can introduce a local définition.

@@

>>>>

\subsection{About  well-foundedness}
\label{sec:T1Wf}
\label{sec:orgheadline82}
   In order to use \texttt{T1} for proving termination results,
we need to prove that  our order \texttt{<} is well-founded. Then we will get \emph{transfinite induction} for free.

@@

\subsubsection{Using a stronger inductive predicate.}
\label{sec:orgheadline78}
\label{sec:strongly-accessible}

  Instead of trying to prove directly that any ordinal term \(\alpha\) in normal form is accessible
through \texttt{LT}, we propose to show first that any well formed 
term of the form \(\omega^\alpha\times(n+1)+\beta\) is accessible (which is a stronger result).

@@

\item Finally, $\canonseq{0}{i}= 0$.
\end{itemize}
\end{mathframe}

@@

\input{movies/snippets/Hprime/HprimeDef}
\input{movies/snippets/Hprime/HprimeDefb}

@@

In order to make this statement more readable, we introduce a local definition.

@@

\input{movies/snippets/Hprime/HprimeOmegaCube3e}
